---
format: gfm
editor: visual
---

# Team Microplastic's FIRE Summit Presentation

Sammi, Neha, Lasya, Sriya, Chayanika

## Research Question

What are the effects of the plastic bag ban on microplastic pollution in the Chesapeake Bay?

## Data Wrangling

**Outcome variable**

Our outcome variable is water quality from each county adjacent to the Chesapeake Bay. Specific water parameters that we are interested in include total nitrogen, total carbon, and total organic nitrogen.

This data is obtained from [https://datahub.chesapeakebay.net](https://datahub.chesapeakebay.net/) which reports water quality from each county each day.

```{r}
#| warning: false
install.packages("tidyverse")
install.packages("terra")
install.packages("simplermarkdown")
install.packages("ggplot2")

library("tidyverse")
library("terra")
library("simplermarkdown")
library("ggplot2")

df<- read.csv("WaterQualityFIPS.csv") %>%
  rename(fips=FIPS)

fips<-read.csv("state_and_county_fips_master.csv")

dfcounty<-merge(df, fips, by="fips", all.x=TRUE) %>%
  mutate(SampleDate=as.Date(SampleDate, format='%m/%d/%Y')) %>%
  mutate(Month=month(SampleDate), Year=year(SampleDate)) %>%
  group_by(Parameter, Unit, fips, name, state, Month, Year) %>%
  summarize(MeasureValue=mean(MeasureValue))
```

**Treatment variable**

Our treatment variable is an indicator of whether there is a plastic bag ban or tax in each county. This data set came from <https://www.bagtheban.com/in-your-state/>.

```{r}
#| warning: false

library("tidyverse")
library("terra")
library("simplermarkdown")

df2<- read.csv("PlasticBagLegislation.csv") %>%
  rename(name=Location) %>%
  rename(state=State) %>%
  mutate(state=ifelse(state=="Maryland", "MD", "VA")) %>%
  mutate(monthBan=ifelse(Ban==1, Month, "N/A")) %>%
  mutate(yearBan=ifelse(Ban==1, Year, "N/A")) %>%
  mutate(monthTax=ifelse(Tax==1, Month, "N/A")) %>%
  mutate(yearTax=ifelse(Tax==1, Year, "N/A"))
```

**Control variables**

To take into account total precipitation and average storm-water runoff data we used data from NASA Earth Data. Our data came from GLDAS Noah Land Surface Model.

## Preliminary Results

```{r}
cmd<-vect("Shapefiles/tl_2020_24_county10.shp")
cva<-vect("Shapefiles/tl_2020_51_county10.shp")

r<-rast("201001.nc4")
rp<-project(r[[1]], crs(cmd))

plot(rp)
plot(cmd, add=TRUE)
plot(cva, add=TRUE)
```

```{r}
df3<- read.csv("PlasticBagLegislation.csv") %>%
  rename(name=Location) %>%
  rename(state=State) %>%
  mutate(state=ifelse(state=="Maryland", "MD", "VA")) 

df4<-df3 %>%
  mutate(monthBan=ifelse(Ban==1, Month, NA)) %>%
  mutate(yearBan=ifelse(Ban==1, Year, NA)) %>%
  mutate(monthTax=ifelse(Tax==1, Month, NA)) %>%
  mutate(yearTax=ifelse(Tax==1, Year, NA)) 

df4<-df4[,-1:-4]

df4county<-merge(df4, dfcounty, by=c("name", "state"), all.y=TRUE) %>%
  mutate(monthBan = ifelse(monthBan=="January", 1, 
                    ifelse(monthBan=="February", 2, 
                    ifelse(monthBan=="March", 3, 
                    ifelse(monthBan=="April", 4, 
                    ifelse(monthBan=="May", 5, 
                    ifelse(monthBan=="June", 6, 
                    ifelse(monthBan=="July", 7,
                    ifelse(monthBan=="August", 8,
                    ifelse(monthBan=="September", 9, 
                    ifelse(monthBan=="October", 10, 
                    ifelse(monthBan=="November", 11, 
                    ifelse(monthBan=="December", 12, NA))))))))))))) %>%
  mutate(monthTax = ifelse(monthTax=="January", 1, 
                    ifelse(monthTax=="February", 2, 
                    ifelse(monthTax=="March", 3, 
                    ifelse(monthTax=="April", 4, 
                    ifelse(monthTax=="May", 5, 
                    ifelse(monthTax=="June", 6, 
                    ifelse(monthTax=="July", 7,
                    ifelse(monthTax=="August", 8,
                    ifelse(monthTax=="September", 9, 
                    ifelse(monthTax=="October", 10, 
                    ifelse(monthTax=="November", 11, 
                    ifelse(monthTax=="December", 12, NA))))))))))))) %>%
  mutate(Ban = ifelse((Year > yearBan), 1, 0)) %>%
  mutate(Ban = ifelse((Year == yearBan & Month>monthBan), 1, Ban)) %>%
  mutate(Tax = ifelse((Year > yearTax), 1, 0)) %>%
  mutate(Tax = ifelse((Year == yearTax & Month>monthTax), 1, Tax))

md_nasa_data <- read.csv("md_nasa_data.csv")
va_nasa_data <- read.csv("va_nasa_data.csv")

#Renaming the names column in order to merge the datasets
md_nasa_data <- md_nasa_data %>%
  rename(name = NAMELSAD10)

va_nasa_data <- va_nasa_data %>%
  rename(name = NAMELSAD10)

# Merging with state_and_county_fips_master to get the state column
md_nasa <- merge(md_nasa_data, fips, by.x = "name")

va_nasa <- merge(va_nasa_data, fips, by.x = "name")

# Creating the month column using the file column
md_nasa <- md_nasa %>%
  mutate(Month = as.numeric(substr((md_nasa$file),5,6)))
va_nasa <- va_nasa %>%
  mutate(Month = as.numeric(substr((va_nasa$file),5,6)))

# Creating the year column
md_nasa <- md_nasa %>%
  mutate(Year = as.numeric(substr((md_nasa$file),1,4)))
va_nasa <- va_nasa %>%
  mutate(Year = as.numeric(substr((va_nasa$file),1,4)))

# Combining va_nasa and md_nasa
all <- rbind(md_nasa, va_nasa)

finaldf <- merge(df4county, all, by=c("name","state", "Month", "Year", "fips"), all.x = TRUE) %>%
  filter(!is.na(Evap_tavg)) %>%
  filter(!is.na(Qs_tavg)) %>%
  filter(!is.na(Rainf_f_tavg)) %>%
  filter(!is.na(Tair_f_tavg))
```

```{r}
# Make a df with the areas with ban, data on before and after water quality(Chesapeake hub bay water)

before_after <- finaldf %>%
  # Getting rid of counties that do not have a ban
  filter(!is.na(monthBan) & !is.na(yearBan)) %>%
  
  #Creating a before column and after column
  mutate(before = ifelse(Year < yearBan & Month < monthBan, finaldf$MeasureValue, NA)) %>%
  mutate(after = ifelse(Year >= yearBan & Month >= monthBan, finaldf$MeasureValue, NA))
  

beforeAverages <- before_after %>%
  filter(!is.na(before)) %>%
  filter(Parameter == "PIC") %>%
  mutate(beforeMean = mean(before))

afterAverages <- before_after %>%
  filter(!is.na(after)) %>%
  filter(Parameter == "PIC") %>%
  mutate(afterMean = mean(after))

dfToGraph <- merge(beforeAverages, afterAverages, by="name", all.x = TRUE)
```
