---
title: "Team Name's FIRE Summit Presentation"
<<<<<<< HEAD
author: "Team Members"
=======
author: "Sammi, Neha, Lasya, Sriya, Chayanika"
>>>>>>> fd548f17025832967e31fc760cb03b2394052205
format: gfm
editor: visual
---

## Research Question

<<<<<<< HEAD
What are the effects of the plastic bag ban on microplastic pollution in the Chesapeake Bay?
=======

>>>>>>> fd548f17025832967e31fc760cb03b2394052205

(Optional) Insert an image if it helps motivate the research question.

## Data Wrangling

**Outcome variable**

<<<<<<< HEAD
Write description, data source, and wrangling methodology. Show a map if you have made one.

If you have loops, do not run the entire thing. Instead, run one sample loop or just display your codes without making it run by adding **#\| eval: false** at the top of the chunk.

You can also suppress warnings by adding **#\| warning: false**.
=======
Our outcome variable is water quality from each county adjacent to the Chesapeake Bay. Specific water parameters that we are interested in include total nitrogen, total carbon, and total organic nitrogen.

This data is obtained from [https://datahub.chesapeakebay.net](https://datahub.chesapeakebay.net/) which reports water quality from each county each day.
>>>>>>> fd548f17025832967e31fc760cb03b2394052205

```{r}
#| warning: false

library("tidyverse")
library("terra")
library("simplermarkdown")
<<<<<<< HEAD
=======

df<- read.csv("WaterQualityFIPS.csv") %>%
  rename(fips=FIPS)

fips<-read.csv("state_and_county_fips_master.csv")

dfcounty<-merge(df, fips, by="fips", all.x=TRUE) %>%
  mutate(SampleDate=as.Date(SampleDate, format='%m/%d/%Y')) %>%
  mutate(Month=month(SampleDate), Year=year(SampleDate)) %>%
  group_by(Parameter, Unit, fips, name, state, Month, Year) %>%
  summarize(MeasureValue=mean(MeasureValue))
>>>>>>> fd548f17025832967e31fc760cb03b2394052205
```

**Treatment variable**

<<<<<<< HEAD
Write description, data source, and wrangling methodology. Show a map if you have one.

**Control variables**

Write description, data source, and wrangling methodology. Show a map if you have one.

## Preliminary Results

Display a figure showing how the treatment variable impacted the outcome variable.
=======
Our treatment variable is an indicator of whether there is a plastic bag ban or tax in each county. This data set came from <https://www.bagtheban.com/in-your-state/>.

**Control variables**
```{r}
cmd<-vect("Shapefiles/tl_2020_24_county10.shp")
cva<-vect("Shapefiles/tl_2020_51_county10.shp")

r<-rast("201001.nc4")
rp<-project(r[[1]], crs(cmd))

plot(rp)
plot(cmd, add=TRUE)
plot(cva, add=TRUE)
```


To take into account total precipitation and average storm-water runoff data we used data from NASA Earth Data. Our data came from GLDAS Noah Land Surface Model.

## Preliminary Results
>>>>>>> fd548f17025832967e31fc760cb03b2394052205


